Estimating the Attack Ratio of Dengue Epidemics under Time-varying Force of Infection using Aggregated Notification Data

### 2015-05-29 Presentations/Dinamica de Vetores-SP-2015/dengueARS0.tex
Updated presentation

\subsection{Variable Force of Infection}
\begin{frame}
\frametitle{Effective Reproductive number(\rr)}
The effective reproductive number can be easily estimated from the incidence 
time-series, $Y_t$:
\begin{equation}
\label{eq:Rtestimate}
R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
\end{equation}
Where $n$ is the ration between the length of reporting interval and 
the mean generation time of the disease.
\begin{flushright}
 Nishiura et. al. (2010)
\end{flushright}
\end{frame}

\begin{frame}
\frametitle{\rr's uncertainty}
But what about the uncertainty about \rr?

If we assume that the counts $Y_t$ are Poisson distributed for all $t$, we can 
derive the probability distribution of \rr\footnote{Coelho, FC and Carvalho, LM 
(Submitted)}
\begin{center}
\begin{figure}
  \includegraphics[width=9cm]{./rt_series.png}
\end{figure}
\end{center}

\subsection{Vector dynamics}
\begin{frame}
\frametitle{Environmental determinants}
\begin{itemize}[<+->]
 \item A. Aegypti population dynamics display marked seasonality
 \item Temperature, Humidity and rainfall are important factors
 \item Environmental stock of eggs
 \item Effects on mosquito reproduction are non-linear
 \item Delayed influence
\end{itemize}

@@

\bibitem{pone2011}
Coelho FC, Carvalho, LM {E}stimating the Attack Ratio of Dengue 
Epidemics under Time-varying Force of Infection using Aggregated Notification 
Data
\newblock Arxiv. \url{http://arxiv.org/abs/1502.01236}
%\bibAnnoteFile{pone2011}
\end{thebibliography}

>>>>

\subsection{Variable Force of Infection}
\begin{frame}
\frametitle{Effective Reproductive number(\rr)}
The effective reproductive number can be easily estimated from the incidence 
time-series, $Y_t$:
\begin{equation}
\label{eq:Rtestimate}
R_t = \left( \frac{Y_{t+1}}{Y_t}\right)^{1/n}
\end{equation}
Where $n$ is the ratio between the length of reporting interval and 
the mean generation time of the disease.
\begin{flushright}
 Nishiura et. al. (2010)
\end{flushright}
\end{frame}

\begin{frame}
\frametitle{\rr's uncertainty}
But what about the uncertainty about \rr\footnote{Coelho, FC and Carvalho, LM 
(Submitted)}?

We explore the approach of Ederer and Mantel\cite{mantel}, whose objective is 
to obtain 
confidence intervals for the ratio of two Poisson counts. 
Let $Y_{t} \sim Poisson(\lambda_t)$ and $Y_{t+1} \sim Poisson(\lambda_{t+1})$ 
and define $S = Y_{t} + Y_{t+1}$.
The authors note that by conditioning on the sum $S$
\begin{align}
\label{eq:binlike}
Y_{t+1} | S &\sim Binomial(S, \theta_t) \\
\theta_t &= \frac{\lambda_{t+1}}{\lambda_{t} + \lambda_{t+1}}
\end{align}
\end{frame}

\begin{frame}
\frametitle{\rr's Uncertainty}
Let $c_{\alpha}(\theta_t) = \{\theta_t^{(L)} , \theta_t^{(U)} \}$ be such that 
$Pr(\theta_t^{(L)}<\theta_t <\theta_t^{(U)}) = \alpha$.
Analogously, define $c_{\alpha}(R_t) = \{R_t^{(L)} , R_t^{(U)} \}$ such that 
$Pr(R_t^{(L)}<R_t<R_t^{(U)}) = \alpha$.

Ederer and Mantel (1974)~\cite{mantel} show that one can construct a $100\alpha 
\%$ confidence interval for \rr by noting that
\begin{equation}
\label{eq:confRt}
 R_t^{(L)} = \frac{\theta_t^{(L)}}{(1-\theta_t^{(L)})} \quad \text{and} \quad 
R_t^{(U)} = \frac{\theta_t^{(U)}}{(1-\theta_t^{(U)})}\\
\end{equation}
\end{frame}

\begin{frame}
\frametitle{\rr's Uncertainty}
Taking a Bayesian conjugate distribution approach, If we choose a
Beta conjugate prior with parameters $a_0$ and $b_0$ for the 
Binomial likelihood in (\ref{eq:binlike}), the posterior distribution for 
$\theta_t$ is
\begin{equation}
\label{eq:thetapost}
p(\theta_t| Y_{t+1}, S) \sim Beta(Y_{t+1} + a_0, Y_t + b_0)
\end{equation}
Combining equations~(\ref{eq:confRt}) and~(\ref{eq:thetapost}) 
tells us that the induced posterior distribution of $R_t$ is 
a Beta prime (or inverted Beta) with parameters $ a_1 = Y_{t+1} + a_0$ and $b_1 
=  Y_t + b_0$~\cite{dubey1970}.
The density of the induced distribution is then 
\begin{equation}
\label{eq:densityMantel}
f_P(R_t| a_1, b_1) = \frac{\Gamma(a_1 + b_1)}{\Gamma(a_1)\Gamma(b_1)} R_t^{a_1 
- 
1} (1 + R_t)^{-(a_1 + b_1)}
\end{equation}
Thus, the expectation of \rr is $a_1/(b_1 - 1)$ and its variance is 
$a_1(a_1 + b_1 - 1)/\left((b_1 - 2)(b_1 - 1)^2 \right) $.
\end{frame}

\begin{frame}
\frametitle{\rr's Uncertainty}
\begin{center}
\begin{figure}
  \includegraphics[width=9cm]{./rt_series.png}
\end{figure}
\end{center}
\end{frame}
\subsection{Vector dynamics}
\begin{frame}
\frametitle{Environmental determinants}
\begin{itemize}[<+->]
 \item A. Aegypti population dynamics display marked seasonality
 \item Temperature, Humidity and rainfall are important factors
 \item Environmental stock of eggs
 \item Effects on mosquito reproduction are non-linear
 \item Delayed influence
\end{itemize}

@@

\bibitem{pone2011}
Coelho FC, Carvalho, LM {E}stimating the Attack Ratio of Dengue 
Epidemics under Time-varying Force of Infection using Aggregated Notification 
Data
\newblock Arxiv. \url{http://arxiv.org/abs/1502.01236}
%\bibAnnoteFile{pone2011}

\bibitem{mantel}
Ederer, F. and Mantel, N. (1974)
\newblock Confidence limits on the ratio of two poisson
  variables.
\newblock \emph{American Journal of Epidemiology}
  \textbf{volume 100:}, pages 165--167

\end{thebibliography}
